# Reproducible Research: Peer Assessment 1

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
```

## Loading and preprocessing the data
```{r, results='markup', warning=TRUE, message=TRUE}
ActivityDS <- read.csv("activity.csv")
```

##### 2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
ActivityDS <- na.omit(ActivityDS)
```

## What is mean total number of steps taken per day?
```{r}
result <- aggregate(ActivityDS$steps, by=list(Category=ActivityDS$date), FUN=sum)
```

##### 1. Make a histogram of the total number of steps taken each day
```{r}
hist(result$x, xlab = "steps", ylab = "day", main = "Total steps per day", col = "blue")
```

##### 2. Calculate and report the mean and median total number of steps taken per day
```{r}
stepsMean <- mean(result$x)
stepsMedian <- median(result$x)
```
* Mean: `r stepsMean`
* Median:  `r stepsMedian`

-----

## What is the average daily activity pattern?
```{r}
result2 <- aggregate(ActivityDS$steps, by=list(Category=ActivityDS$interval), FUN=mean)

```

##### 1. Make a time series plot
```{r}
plot(result2$Category, result2$x, type="l", main = "Average Daily Activity Pattern", xlab="Steps", ylab="Interval")
```

##### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxSteps <- subset(result2, result2$x == max(result2$x))
```

* Max Steps: `r maxSteps`

----

## Inputing missing values
##### 1. Calculate and report the total number of missing values in the dataset 
```{r}
ActivityDS <- read.csv("activity.csv")
missingValuesTotal <- sum(is.na(ActivityDS))
```

* Number of missing values: `r missingValuesTotal`

##### 2. Devise a strategy for filling in all of the missing values in the dataset.
##### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
NApositions <- which(is.na(ActivityDS$steps))
meanVector <- rep(mean(ActivityDS$steps, na.rm = T), times = length(NApositions))
ActivityDS[NApositions, "steps"] <- meanVector
head(ActivityDS)
```


##### 4. Make a histogram of the total number of steps taken each day 
```{r}
result3 <- aggregate(ActivityDS$steps, by=list(Category=ActivityDS$date), FUN=sum)
hist(result3$x, xlab = "steps", ylab = "day", main = "Total steps per day", col = "blue")
```

##### ... and Calculate and report the mean and median total number of steps taken per day. 
```{r}
stepsNewMean <- mean(result3$x)
stepsNewMedian <- median(result3$x)
```
* Mean (Imputed): `r stepsNewMean`
* Median (Imputed):  `r stepsNewMedian`


----

## Are there differences in activity patterns between weekdays and weekends?
##### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
weekDay <- function(date) {
  day <- weekdays(date)
  # Due to my locale all days are written in Brazilian Portuguese
  if (day %in% c("segunda", "terça", "quarta", "quinta", "sexta"))
    return("weekday")
  else if (day %in% c("sábado", "domingo"))
    return("weekend")
  else
    stop("no valid date")
}

ActivityDS$date <- as.Date(ActivityDS$date)
ActivityDS$day <- sapply(ActivityDS$date, FUN = weekDay)
```

##### 2. Make a panel plot containing a time series plot

```{r}
average <- aggregate(steps ~ interval + day, data=ActivityDS, mean)
ggplot(average, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) + xlab("Interval") + ylab("Number of Steps")
```

